### 2020-10-17 Thesis-Chapter1-Introduction.tex
Update OODA model and begin writing chap 3

Certain policy issues are \acp{wicked-prob}. Originally coming from the field of design theory, the term
``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
\cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science and engineering are well equipped to
answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or measurements. Each attempted solution has
permanent and often unintended consequences, and likely exists in a pattern of chronic policy failure.

>>>>

Certain policy issues are \acp{wicked-prob}. Originally coming from the field of design theory, the term
``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
\cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science, engineering, and traditional policymaking
are well equipped to answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or measurements. Each
attempted solution has permanent and often unintended consequences, and likely exists in a pattern of chronic policy
failure.

### 2020-10-17 Thesis-Chapter3-Arguments.tex
Update OODA model and begin writing chap 3

Before going into detail about the arguments used in the \ac{EA} debate, it is worth studying \acp{wicked-prob} in
greater detail. What are they, and why is the case of encryption, privacy, and \ac{EA} in this category? What approaches
to tackling \acp{wicked-prob} succeed and fail? How can we collectively situation ourselves to make real progress? This
chapter seeks to answer these questions.

@@

@@

>>>>

Before going into detail about the arguments used in the \ac{EA} debate, it is worth studying \acp{wicked-prob} in
greater detail. What are they, and why is the case of encryption, privacy, and \ac{EA} in this category? What approaches
to tackling \acp{wicked-prob} succeed and fail? How can we collectively situate ourselves to make real progress? This
chapter seeks to answer these questions.

@@

\Acp{wicked-prob} were previously introduced as pernicious and tricky issues that resist straightforward solutions.
In this section we analyze the nature of wicked problems and strategies for tackling them.

@@

Rittel's categorization of \acp{wicked-prob} grew out of frustration with their resistance to classical problem solving
methods. Since the Enlightenment, society has applied the scientific method to problems of every kind; the sweeping
application of scientific analysis has delivered reliable clean water, improved crop yields, shaped government
structures, and bestowed material wealth previously unimaginable. With our material problems largely solved in the
twentieth century, believers in the power of reason thought this progress would march on in the realm of public
planning. Policymaking would function by setting goals, identifying problems, evaluating alternatives, implementing
solutions, and analyzing outcomes in order to correct errors. Functioning as a continuous process, this approach was
primed to revolutionize governing the same way it did industry, agriculture, and economics---until it didn't. In the
context of what he describes as an ``anti-professional movement,'' Rittel explains how the scientific method has failed:

\begin{displayquote}
A great many barriers keep us from perfecting such a planning/governing system: theory is inadequate for decent
forecasting; our intelligence is insufficient to our tasks; plurality of objectives held by pluralities of politics
makes it impossible to pursue unitary aims; and so on. The difficulties attached to rationality are tenacious, and we
have so far been unable to get untangled from their web. This is partly because the classical paradigm of science and
engineering---the paradigm that has underlain modern professionalism---is not applicable to the problems of open
societal systems. One reason the publics have been attacking the social professions, we believe, is that the cognitive
and occupational styles of the professions---mimicking the cognitive style of science and the occupational style of
engineering---have just not worked on a wide array of social problems. The lay customers are complaining because
planners and other professionals have not succeeded in solving the problems they claimed they could solve. We shall want
to suggest that the social professions were misled somewhere along the line into assuming they could be applied
scientists---that they could solve problems in the ways scientists can solve their sorts of problems. The error has been
a serious one. \cite{rittel_dilemmas_1973}
\end{displayquote}

When applied to social problems, the prescribed method---here defined as setting goals, identifying problems, evaluating
alternatives, implementing solutions, and analyzing outcomes---fails at every step. In the U.S., a nation with countless
cultures and subcultures which are effectively represented by two mutually hostile political parties, agreeing on goals
is a challenge in itself. When goals are set, we often discover that we are contending with \acp{wicked-prob} that defy
the process at each remaining step. Rittel provides a list of ten characteristics of problems in this category
\cite{rittel_dilemmas_1973}:

\begin{displayquote}
  \begin{enumerate}
    \item There is no definitive formulation of a wicked problem [and each formulation presupposes a solution].
    \item Wicked problems have no stopping rule.
    \item Solutions to wicked problems are not true-or-false, but good-or-bad.
    \item There is no immediate and no ultimate test of a solution to a wicked problem.
    \item Every solution to a wicked problem is a ``one-shot operation''; because there is no opportunity to learn by
          trial-and-error, every attempt counts significantly.
    \item Wicked problems do not have an enumerable (or an exhaustively describable) set of potential solutions, nor is
          there a well-described set of permissible operations that may be incorporated into the plan.
    \item Every wicked problem is essentially unique.
    \item Every wicked problem can be considered to be a symptom of another problem.
    \item The existence of a discrepancy representing a wicked problem can be explained in numerous ways. The choice of
          explanation determines the nature of the problem's resolution.
    \item The planner has no right to be wrong [i.e., they are liable for the consequences of their decisions].
  \end{enumerate}
\end{displayquote}

Based on this description, \acp{wicked-prob} are impossible to definitively identify, their alternative solutions are
infinite in number but cannot be proved or independently tested, solution implementations carry large risks, and outcome
analysis is subject to interpretation. The policy version of the scientific method cannot be used under these
conditions.

Several of these characteristics are results of the fact that we have no accurate predictive model of the world and
human behavior at large. This does not need to be expounded upon.

The more interesting insight is that problem definition is not self-contained. There is no way to frame the problem that
provides all the information required for a solution...

% Cite commission_tackling_2018 and some other wicked problems
